\documentclass[12pt,english]{article}

\usepackage{amsmath}
\usepackage[margin=1in]{geometry}
\begin{document}

%\vspace{-4cm}
\begin{center}
{\large Exercise 2} \\
The Riemann problem for acoustic waves at a material interface
\par\end{center}{\large \par}

Consider the Riemann problem given by the variable-coefficient
acoustics equations
\begin{align}
p_t + K(x) u_x & = 0 \\
u_t + \frac{1}{\rho(x)} p_x & = 0
\end{align}
with
\begin{align}
(\rho(x), K(x)) & = \begin{cases} (\rho_\ell, K_\ell) & x < 0 \\ (\rho_r,K_r) & x > 0\end{cases} \\
(p(x,0),u(x,0)) & = \begin{cases} (p_\ell, u_\ell) & x < 0 \\ (p_r,u_r) & x > 0.\end{cases}
\end{align}
In this exercise we will find the solution to this problem and in the process rediscover
the classical transmission and reflection coefficients.

(a)  The above system can be written as $q_t + A(x)q_x = 0$.
Show that for any real, positive values $K(x), \rho(x)$, the matrix $A(x)$ has
one positive and one negative eigenvalue, so that the system is hyperbolic with
one family of waves going in each direction.

Notice that we cannot define a single set of characteristic variables, since they depend
on the coefficients that vary in space.

(b)  The Riemann solution will consist of a wave (or discontinuity) going to the left in the medium
with coefficients $(\rho_\ell, K_\ell)$ and a wave (or discontinuity) going to the right
in the medium with coefficients $(\rho_r, K_r)$.  To find this solution we must decompose the
jump $q_r - q_\ell$ into a linear combination of the corresponding eigenvectors of the matrices
$A(x<0)$ and $A(x>0)$.  That is, we must use the eigenvector corresponding to left-going waves
in the left medium and the eigenvector corresponding to right-going waves in the right medium.
Write down and solve this linear system of equations.

(c) Now suppose that the initial data corresponds to a jump that is incident from the left;
i.e., that $q_r - q_\ell$ is a multiple of the left-going wave family eigenvector of the left medium.
For instance, you could take $q_r = [0,0]^T$ and $q_\ell = [-Z_\ell, 1]^T$.  Compute the solution of
the Riemann problem, and then find:

\begin{itemize}
    \item The ratio of the left-going wave to the initial jump; this is the reflection coefficient.
    \item The ratio of the right-going wave to the initial jump; this is the transmission coefficient.
\end{itemize}

\end{document}
